Disc stuff

### 2012-10-23 paper.tex
Revert "stuff"
This reverts commit 1670e6e4081bbc62f0d8bab3b33c2d541a565556.

The model used to simulate the emission from a young protoplantary disc is a hybrid model created from hydrodynamical models of an infalling prestellar core and a gravitationally unstable protoplanetary disc. The prestellar core model is based upon the model of Keto \& Caselli (2010). This describes an infalling pre-stellar core, with densities temperatures and velocities given by the model of the collapse of a 10 solar mass Bonnor-Ebert sphere and provides a match to line profiles seen in the pre-stellar core L1544. Some slight modification due to the addition of Oxygen lines (Keto et al 2012 in prep) have been been added to this model resulting in a slight decrease in the peak infall velocity. This prestellar core model extends from 80$\,$au out to 10,000$\,$AU. The model is a 1D spherically symmetric model with inward motions (see figure \ref{l1544model}).

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/L1544model_used.png}
 \label{l1544model}
 \caption{envelope model showing temp and dust temp (red, green) in kelvin, log number density (blue) in cm$^{-3}$ and inward velocity$\times$10 (cyan) in m$\,$s$^{-1}$, adapted from Keto \& Caselli (2010)}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/rhoT.png}
 \label{discModel}
 \caption{The density and temperatures of the 0.4$\,\rm{M}_\odot$ gravitationally unstable disc model used in inside 80$\,$AU. The dense and hot regions are confinded to a few AU about the rotation plane.}
\end{figure*}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/velocity_rzSlice_thetaColourScale2.png}
 \label{discVel}
 \caption{The line of sight (colour scale) and plane of sky (arrows) velocities in the y=0 slice through the disc model.}
\end{figure*}

At radii less than 80$\,$au the physical structure is the given by the hydrodynamic model described in Ilee et al. (2011), based on the work of Boley(2007), and Boley(2009). This model describes a 0.39$\,\rm{M}_\odot$ self-gravitating disc featuring prominent spiral arms. H$_2$ number densities in the disc range from 10$^{10}$-10$^{19}\,\rm{m}^{-3}$, and temperatures range from 30-400 K. The dust and gas temperatures in the disc are assumed to be in equilibrium. The model is sampled over a regular grid of size 256$\times$256$\times$64 with spatial resolution of 0.5$\,$au in x, y and z. The majority of the mass lies in the mid-plane of the disc. For the models using a smoothed disc the same physical and chemical model is used but with the temperature, density and abundance averaged in $\phi$. The gas/dust mass ratio was assumed to be 1/100 throughout both sections model and the opacities were given by the model of dust grains with thick icy mantles and 10$^6\,$yr coagulation from Ossenkopf and Henning (1994)\newline

Chemical abundances in the disc were taken from Ilee et. al (2011) which followed the changes of chemical abundances of trace particles moving through the disc as it evolved. The abundances of 125 species related by 1334 reactions were followed through the time evolution of the disc. These abundances were interpolated onto a 51$^3$ grid covering the disc with cells of size 2.2$\times$2.2$\times$0.22$\,$au. Abundances in the envelope model were as follows:\newline

(tableify this?)\newline
HCO$^+$: as the H2O profile from (find out where L1544 H2O profile is from) scaled so that the maximum abundance is 10$^{-8}$\newline
HNO: constant abundance of 5$\times$10${-11}$\newline
HCS$^+$: 10$^{-11}$\newline
OCS: constant abundance of 10$^{-9}$\newline
H$_2$CO: 1.5$\times$10$^{-9}$ decreased by a factor of 40 at radii less than 8250au (Young et al 2004)\newline
CS:  3$\times$10$^{-9}$ decreased by a factor of 10,000 at radii less than 6700au (Tafalla, santiago-garcia and myers 2006)\newline
CO: as the H2O profile from before scaled so that the maximum abundance is CO maximum\newline

\section{description of radiative transfer program}
LIME (Brinch 2011), the radiative transfer program used, calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3d model. The method of selecting these points is given in the griding section. At each of these points the density of the main collision partner (in this case H$_2$), gas and dust temperatures, velocity, molecular abundances and turbulent velocity and taken from the model. These points are then smoothed by Lloyds algorithm (Lloyd 1982) in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is down these paths that photons are restricted to propagating (figures \ref{grid}). The levels of molecules in question are then  calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value.\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}
 \label{grid}
 \caption{An example of the points distribution chosen by lime for a model of a smooth protoplanetary disc. The nodes of the grid are the points at which the underlying physical model is sampled and the edges are teh paths down which photons can travel during the excitation-radiation emission stage of the computation.}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/lime_points_rz2.png}
 \label{points}
 \caption{the rz distribution of points selected to sample the model. The points in the dense region at 10$^9<\rm{z}<10^12$ and 7*10$^{11}<\rm{r}<10^{13}$ are sampling the disc model and the points at larger values are those foreced into the outer regions to sample the envelople.}
\end{figure}

In order to construct the grid, points are randomly selected from the volume being simulated then compared against a reference point. Grid points are selected at random in cylindrical co-ordinates, linearly spaced in z and $\phi$ and logarithmically spaced in r. For each point to be selected a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random co-ordinates the Hydrogen density and molecular density at the point (n and m) are compared against the densities of a reference point on the inner edge of the disc (n$_0$ and m$_0$). If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use, if not then another r, $\phi$, z co-ordinate is selected. The weighting function and griding functions were selected empirically to sample both the all scales while ensuring the majority of the points went into the inner disc which is the region of interest. 20\% of these points are forced to be at radii greater than $\sqrt{R_{min}R_{max}}$ (where $R_{min}$ and $R_{max}$ are the inner and outer radius of the model) in order to stop too many of the selecected points clustering in the high density disc and leaving the envelope undersampled. In addiation to this method of selection 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. See figure \ref{points} for example of the points distribution in r, z. \newline

@@

Continuum emission from the model. Able to do this as the majority of the computing time on the model is taken up with calculating molecule population levels. This image has been created by averaging the output images of 100 unique runs. This averaging mitigates the granularity of output images from lime due to the nature of the gridding.

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.png}

 \caption{A composite image of 100 runs of continuum emission at 350$\,$GHz}
\end{figure}

>>>>

-describe the envelope structure (pre-stellar core)
Keto \& Caselli model (2010), disc is embedded in an infalling pre-stellar core, with densities temperatures and velocities given by the model of the collapse of a 10 solar mass Bonnor-Ebert sphere and providing similar line profiles to the pre-stellar core L1544 as described by Keto \& Caselli (2010), with some slight modification due to the addition of Oxygen lines (keto et al 2012 in prep). This prestellar core model extends from 80$\,$au out to 10,000$\,$au. The model is 1D spherically symmetric model with inward motions (see figure).

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/L1544model_used.png}

 \caption{envelope model with temp and dust temp (red, green) in kelvin, log number density (blue) in cm$^{-3}$ and inward velocity$\times$100 (cyan) in m$\,$s$^{-1}$}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/rhoT.png}

 \caption{this is a density \& temp plot}
\end{figure*}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/velocity_rzSlice_thetaColourScale2.png}

 \caption{this is a velocity slice in the rz plane}
\end{figure*}

-describe the physical structure of the disk (density and temperature) -- one figure showing the physical structure and the kinematics (2-panels figure)
The physical structure of the disc is the same as the one used in Ilee et al. (2011), based on the work of Boley(2007), Boley \& Durisen(2008) and Boley(2009). This model describes a 0.39$\,\rm{M}_\odot$ self-gravitating disc featuring prominent spiral arms. Densities in the disc range from 10$^{10}$-10$^{21}\,\rm{m}^{-3}$, and temperatures range from 30-400 K. The dust and gas temperatures in the disc are assumed to be equal. The gas/dust mass ratio was assumed to be 1/100 throughout the model and the opacities were given by the model of dust grains with thick icy mantles and 10$^6\,$yr coagulation from Ossenkopf and Henning (1994). The model is sampled over a regular grid of size 256$\times$256$\times$64 with spatial resolution of 0.5$\,$au in x, y and z. The majority of the mass lies in the mid-plane of the disc. For the models using a smoothed disc the same physical and chemical model is used but with the temperature, density and abundance averaged in $\phi$.\newline

-describe the chemistry  - refer to Ilee et al. which have been taken as input to the rad transfer (RT) code
Chemical abundances in the disc were taken from Ilee et. al (2011) which followed the changes of chemical abundances of trace particles moving through the disc as it evolved. The abundances of 125 species related by 1334 reactions were followed through the time evolution of the disc. These abundances were interpolated onto a 51$^3$ grid covering the disc with cells of size 2.2$\times$2.2$\times$0.22$\,$au. Abundances in the envelope model were as follows:\newline

(tableify this?)\newline
HCO$^+$: as the H2O profile from (find out where L1544 H2O profile is from) scaled so that the maximum abundance is 10$^{-8}$\newline
HNO: constant abundance of 5$\times$10${-11}$\newline
HCS$^+$: 10$^{-11}$\newline
OCS: constant abundance of 10$^{-9}$\newline
H$_2$CO: 1.5$\times$10$^{-9}$ decreased by a factor of 40 at radii less than 8250au (Young et al 2004)\newline
CS:  3$\times$10$^{-9}$ decreased by a factor of 10,000 at radii less than 6700au (Tafalla, santiago-garcia and myers 2006)\newline
CO: as the H2O profile from ??? scaled so that the maximum abundance is CO maximum\newline\

-describe the RT used (LIME) 
LIME (Brinch 2011), the radiative transfer program used, calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3d model. The method of selecting these points is given in the griding section. At each of these points the density of the main collision partner (in this case H$_2$), gas and dust temperatures, velocity, molecular abundances and turbulent velocity and taken from the model. These points are then smoothed by Lloyds algorithm (Lloyd 1982) in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is down these paths that photons are restricted to propagating. The levels of molecules in question are then  calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. \newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}

 \caption{plot of the griding overlaid on a smoothed density model}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/lime_points_rz2.png}

 \caption{plot of the points selected in the model}
\end{figure}

In order to construct the grid, points are randomly selected from the volume being simulated then compared against a reference point. Grid points are selected at random in cylindrical co-ordinates, linearly spaced in z and $\phi$ and logarithmically spaced in r. For each point to be selected a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random co-ordinates the Hydrogen density and molecular density at the point (n and m) are compared against the densities of a reference point on the inner edge of the disc (n$_0$ and m$_0$). If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use, if not then another r,$\phi$,z co-ordinate is selected. The weighting function and griding functions were selected empirically to sample both the all scales while ensuring the majority of the points went into the inner disc which is the region of interest. 20\% of these points are forced to be at radii greater than $\sqrt{R_{min}R_{max}}$ (where $R_{min}$ and $R_{max}$ are the inner and outer radius of the model) in order to stop too many of the selecected points clustering in the high density disc and leaving the envelope undersampled. In addiation to this method of selection 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. See figure 5 for example of the points distribution in r, z.

@@
